\chapter{RESULTS}

\section{In Silico Ground Truth Extraction and Hardware Benchmarking}
Prior to empirical applications, the NECTA mathematical modules were tested \textit{in silico} against synthetic data matrices. By feeding the architecture with pre-programmed neural mass signals embedded in progressive layers of Pink/White noise and modeled electrical spatial leakage (simulated Volume Conduction), the `wPLI` and dynamically optimized PDC estimators consistently filtered the spurious zero-phase interactions. Analyzing the Receiver Operating Characteristic (ROC) outputs confirmed high precision in extracting the embedded causal ground-truth graphs, despite severe Signal-to-Noise (SNR) corruption.

Concurrently, hardware stress testing evaluated the distributed execution of these complex tasks. The monolithic extraction of MVAR matrices combined with phase-randomized surrogate networks (e.g., iterating through multiple frequency bands) resulted in deterministic Out-of-Memory (OOM) failures or intractable execution times when running sequentially. However, by deploying NECTA’s Dask-driven orchestrator, linear scalability was achieved across active worker threads (1, 2, 4, 8, 16), effectively accelerating combinatorial FDR validation and dynamically preventing memory exhaustion by leveraging chunked Zarr abstractions.

\section{Baseline Characterization and the Static Backbone}
To establish a baseline structural framework, static functional connectivity was evaluated across the entire recording session. Applying Magnitude Coherence in the Low Gamma band (30–59 Hz) with a proportional density threshold of 15%, the NECTA framework successfully reconstructed the macro-architecture of the recorded visual areas. The Louvain community algorithm blindly extracted functional clusters that mapped perfectly onto the physical anatomical placement of the multielectrode arrays, naturally partitioning into three distinct topological modules corresponding to Area 17 (A17), the Transition Zone (TZ), and Area 18 (A18). This static connectome serves as the anatomical "structural backbone" \cite{vezoli2021}.

\section{Hierarchical Information Flow (Directed Connectivity)}
To decode the causal relationships within this structural backbone, Partial Directed Coherence (PDC) was applied using the BIC-optimized MVAR models. The static PDC analysis in the Low Gamma band revealed a highly asymmetrical and hierarchical information flow across the visual cortex. 
By calculating the Net Flow for each region, A17 emerged as the dominant persistent \textit{Driver} (Source) of the network (+1.10 net flow). In contrast, A18 functioned as the primary \textit{Receiver} (Sink) (-1.36 net flow). These results establish a clear directional pathway extending from A17 through the TZ and terminating in A18 (A17 \&rarr; TZ \&rarr; A18), consistent with canonical feedforward processing hierarchies.

\section{Dynamic Topology and Temporal Integration}
The core capability of NECTA lies in capturing the "Dynome"—the transition from static networks to chronological connectomes. Connectivity was recalculated using a sliding-window approach ($W = 500$ ms, 50% overlap). The Temporal Integration panel overlaid the global population spiking activity (Peri-Stimulus Time Histogram - PSTH) with the topological evolution of the network's \textit{Small-Worldness} ($\sigma$).

The dynamic analysis revealed that the network's topological efficiency is not static but responds organically to visual processing phases. The network achieved its absolute peak in \textit{Small-Worldness} specifically during the moving grating stimulus phase. This transient optimization was driven primarily by a rapid surge in local clustering, indicating that the anesthetized cortical network actively reorganizes into a highly efficient state to process dynamic visual inputs.

\vspace{1cm}

% --- VISUAL INSERTION 3 USING TIKZ ---
\begin{figure}[htbp]
    \centering
    \begin{tikzpicture}[
        >=Stealth,
        node distance=3.5cm,
        area/.style={circle, draw, minimum size=1.5cm, font=\sffamily\bfseries, thick, align=center},
        source/.style={fill=red!20, draw=red!80!black},
        sink/.style={fill=blue!20, draw=blue!80!black},
        mod/.style={fill=purple!20, draw=purple!80!black},
        neutral/.style={fill=gray!10, draw=gray},
        strongflow/.style={->, ultra thick, draw=red!70!black},
        weakflow/.style={->, thick, dashed, draw=gray},
        activeflow/.style={->, line width=2.5pt, draw=orange!90!black}
    ]

    % --- PANEL A: Baseline ---
    \node[font=\sffamily\bfseries] at (0, 3) {Painel A: Período Basal (Pré-Estímulo Silencioso)};
    
    \node[area, source] (A17_A) at (-3, 0) {A17\\(+ Net)};
    \node[area, neutral] (TZ_A) at (0, -2) {TZ\\(Neutro)};
    \node[area, sink] (A18_A) at (3, 0) {A18\\(- Net)};
    
    % Flows for Panel A
    \draw[strongflow] (A17_A) -- node[above, font=\sffamily\scriptsize] {Driver Dominante} (A18_A);
    \draw[weakflow] (A17_A) -- (TZ_A);
    \draw[weakflow] (TZ_A) -- (A18_A);
    
    \node[font=\sffamily\scriptsize, text=gray!80!black, align=center] at (0, -3.2) {Fluxo Crônico em Repouso:\\ Hierarquia A17 $\rightarrow$ A18};

    % --- PANEL B: Stimulus Peak ---
    \begin{scope}[yshift=-7.5cm]
        \node[font=\sffamily\bfseries] at (0, 3) {Painel B: Pico Modular (Entrada do \textit{Grating})};
        
        \node[area, source] (A17_B) at (-3, 0) {A17\\(Fonte)};
        \node[area, mod, fill=orange!30, draw=orange!80!black] (TZ_B) at (0, -2) {TZ\\(Ativa)};
        \node[area, sink] (A18_B) at (3, 0) {A18\\(Sumidouro)};
        
        % Flows for Panel B
        \draw[strongflow, line width=1.5pt] (A17_B) -- (A18_B);
        \draw[activeflow] (TZ_B) -- node[left=0.2cm, font=\sffamily\scriptsize, align=right] {Inversão\\Direcional} (A17_B);
        \draw[activeflow] (TZ_B) -- node[right=0.2cm, font=\sffamily\scriptsize, align=left] {Modulação\\Intensa} (A18_B);
        
        \node[font=\sffamily\scriptsize, text=orange!80!black, align=center] at (0, -3.2) {Polaridade Mutável:\\ TZ coordena a integração espacial};
    \end{scope}

    \end{tikzpicture}
    \caption[Mapeamento Evolutivo de Net Flow]{\textbf{Mapeamento Evolutivo Topológico do Net-Flow de Informação (Banda Baixa-Gama).} Proporções de entrada (\textit{in-degree}) menos saída (\textit{out-degree}). \textbf{(A)} Durante o período basal, a rede opera numa hierarquia visual preditiva estrita, onde a Área 17 (A17) é o polo motriz primário e a Área 18 (A18) o sumidouro passivo, com a Zona de Transição (TZ) em estado neutro. \textbf{(B)} No exato momento do estímulo (inserção do \textit{grating} móvel), a topologia sofre uma remodelação ativa. A TZ reverte a sua polaridade passiva e adota dinâmicas de modulação aguda, alterando drasticamente os \textit{edge weights} (espessura das setas) e provando que a flexibilidade mesoscópica reage transitoriamente à entrada do estímulo periférico.}
    \label{fig:netflow_evolution}
\end{figure}

\vspace{0.5cm}

\section{The "Dynome" Evolution and Modulatory Nodes}
Extracting frame-by-frame snapshots (Connectome Filmstrip) of the dynamic Low Gamma Coherence network illustrated a stable rhythmic organization over time. The hubs located in A17 and the TZ maintained cohesive clustering. Crucially, evaluating the \textit{Net Flow Evolution} uncovered a complex temporal interplay obscured by static analysis. While A17 consistently acted as a "Volatile Source" and A18 as a "Stable Sink", the Transition Zone (TZ) exhibited unique behavior. The TZ functioned as a \textit{Dynamic Modulator}, actively switching its directional role (from sink to source, or vice versa) precisely at the onset of visual stimulation. This highlights the framework's capacity to dissect the mesoconnectome chronologically, revealing dynamic hierarchical relays that coordinate sensory integration across the interhemispheric boundary.