\chapter{DISCUSSION}

\section{From Static Maps to Dynamic Visual Networks}
Historically, the functional organization of the primary visual system has been delineated through the static mapping of receptive fields and anatomical inter-hemispheric tracing \cite{wunderle2013}; \cite{schmidt2023}. While these classical methods successfully outline the spatial constraints and functional columns of the cortex, they inherently capture only a time-averaged "structural backbone." Cognitive processing and the integration of moving stimuli, however, depend on millisecond-scale state transitions within neural ensembles—a temporal dimension of rapid network evolution \cite{kopell2014}. 

Even as our anatomical maps reach unprecedented levels of precision---such as recent efforts updating the classic Binzegger cortical wiring diagram to reveal massive, previously unaccounted local excitatory synapses terminating in layer 6 \cite{Martin_2024}---structure alone cannot predict traffic. While anatomy dictates the available pathways, only time-resolved effective connectivity frameworks can reveal how this intricate wiring operates on a millisecond scale.

The application of the NECTA framework to the \texttt{c1608a01} dataset provides empirical proof of this dynamic nature. By utilizing a continuous sliding-window approach, we demonstrated that the visual cortex undergoes rapid, stimulus-locked topological reorganizations. Furthermore, recent \textit{in vitro} microfluidic studies by \cite{Lassers2026} have demonstrated that theta oscillations (4--10 Hz) in isolated axons occur spontaneously and independently of spiking, actively coordinating target neuron bursting based on their phase and amplitude. Strikingly, this axonal transmission relies on a sparse coding architecture, where only a restricted fraction of axons carries high-amplitude signals. This cellular mechanism perfectly underpins our mesoscopic observation that the network reaches transient peaks in \textit{Small-Worldness} precisely during the moving grating stimulus. The transient optimization detected by NECTA is likely mediated by this selective, sparse recruitment of communication channels, demonstrating that macroscopic network efficiency is an active, responsive property rather than a fixed anatomical trait.

This transient optimization is strongly supported by recent anatomical discoveries. \cite{Martin_2024} identified an extraordinarily extensive horizontal spread within Area 17, with neurons integrating information across distances of up to 11 millimeters (spanning 5-10 degrees of the visual field). This vast structural lateral arborization provides the exact physical machinery required for the network to rapidly bind global spatial information, perfectly explaining the acute surges in \textit{Small-Worldness} we detect computationally. Furthermore, their discovery of a novel, direct ``shortcut'' projection from the layer 3/4 border to layer 6---bypassing the canonical layer 5 relay and likely associated with the fast-conducting, motion-sensitive Y-pathway---underscores the necessity of high-frequency dynamic analyses. The existence of such fast parallel anatomical routes proves that visual processing is not strictly serial or slow, fully justifying our focus on the gamma band and sliding-window techniques to capture these rapid, transient synchronizations that static tools completely miss. The transient optimization detected by NECTA is likely mediated by this selective, sparse recruitment of communication channels, demonstrating that macroscopic network efficiency is an active, responsive property rather than a fixed anatomical trait.

\section{Isolating Intrinsic Dynamics in the Anesthetized State}
A critical strength of our biological validation lies in the physiological state of the animal model. The recordings were obtained under general anesthesia and complete paralysis. Far from being a limitation, this state is analytically profound. It systematically suppresses top-down cognitive and attentional modulations, behavioral volition, and saccadic eye movements. 
Therefore, the dynamic topological reorganizations and the causal directionality shifts extracted by NECTA—such as the transient peaks in network clustering—cannot be dismissed as mere artifacts of cognitive states or conscious attentional engagement. Instead, the framework successfully unveils the intrinsic, low-level reflexive plasticity inherent to local synaptic loops and dense transcallosal pathways interlinking A17, A18, and the TZ. Revealing this level of structural fluidity without massive executive-level interference highlights NECTA’s power as a tool for decoding primary mesoscopic coding.

\section{Mitigating Electrophysiological Confounders}
A profound challenge in high-density LFP connectomics is the pervasive artifact of volume conduction, wherein a single electrical source is instantaneously registered by multiple neighboring electrodes \cite{bastos2015}. If unaccounted for, this phenomenon generates spurious zero-phase lag correlations, heavily biasing the resulting graph topology with false-positive edges. 

NECTA addresses this methodological bottleneck architecturally. By employing the Hilbert Transform to extract analytical signals, the framework calculates phase-based indices such as the Weighted Phase Lag Index (\textit{wPLI}) and Imaginary Coherence (\textit{iCoh}), which are mathematically immune to instantaneous spatial spread. This algorithmic refinement ensures that the dynamic topologies extracted reflect true underlying physiological synchronization rather than electrical crosstalk. This distinction is critically reinforced by recent experimental evidence isolating single axons in microfluidic tunnels, which proved that subthreshold voltage oscillations are genuine, regenerative local phenomena of the axon rather than mere volume-conducted epiphenomena of the extracellular matrix \cite{Lassers2026}. By utilizing \textit{wPLI} and \textit{iCoh}, NECTA's architecture is perfectly aligned with this biological reality, mathematically isolating true phase-delayed interactions.

\section{Gamma Synchronization, Predictive Coding, and Callosal Modulation}
Previous work from our group has established the cardinal role of gamma-band synchronization in functional routing and cohesion within the retinogeniculate and cortical pathways of the cat \cite{neuenschwander2023}. NECTA extends this foundation by mapping the causal directionality of these high-frequency rhythms. 

By applying Partial Directed Coherence (PDC) optimized dynamically via the Bayesian Information Criterion (BIC), we eliminated the bias of static model-order selection (which often leads to overfitting when estimating temporal shifts, unlike AIC). The resulting hierarchy in the Low Gamma band (A17 acting as the persistent Driver, A18 as the Receiver) strongly corroborates current theories of predictive coding and feedforward information routing \cite{vezoli2021}. At the cellular level, this macroscopic directionality is supported by findings that the phase of axonal theta oscillations dictates the firing probability of target neurons, establishing a clear physical causality where subthreshold oscillations `prime' target neurons for spiking \cite{Lassers2026}. 

The validity of utilizing PDC for this structural mapping is strongly supported by recent large-scale computational studies. Nunes et al. (2021) demonstrated \textit{in silico} using spiking neuronal networks that generalized PDC provides a highly reliable estimate of true underlying anatomical structural connectivity (yielding a Pearson correlation of $r=0.74$). Crucially, their model intrinsically generated LFP signals with prominent spectral peaks in the gamma band, directly validating our empirical focus on low-gamma (30--59 Hz) synchronization as a primary channel for causal routing. Furthermore, a persistent challenge in high-density electrophysiology is the ``partial observation'' problem, where unrecorded brain regions might introduce spurious causalities. \cite{Nunes_2021} proved that PDC estimates maintain robust correlations with structural connectivity ($r > 0.6$) even when the analysis is restricted to a small spatial cluster. This provides powerful theoretical justification that the causal inferences drawn by NECTA from the restricted A17-TZ-A18 microcircuit remain statistically valid despite the absence of whole-brain recordings. At the cellular level, this macroscopic directionality is supported by findings that the phase of axonal theta oscillations dictates the firing probability of target neurons, establishing a clear physical causality where subthreshold oscillations `prime' target neurons for spiking \cite{Lassers2026}. 

Furthermore, identifying the Transition Zone (TZ) as a "Dynamic Modulator" that reverses its causal flow upon stimulus onset highlights the necessity of time-resolved directional tools. Anatomically, the TZ boundary is densely innervated by mixed intra-hemispheric projections and transcallosal fibers that align spatial orientation maps across the vertical meridian \cite{schmidt2023}. By switching from a passive sink to an active causal source during the moving grating stimulation, the TZ likely orchestrates the interhemispheric temporal unifications required to track sensory borders crossing visual fields. Static connectivity would have merely averaged this reversal into an indiscernible null-flow, masking a critical operational mechanism of visual processing. This extreme functional volatility is theoretically supported by the concept of ``activity flow'' across neural networks, where the variability of a node's directed functional connectivity is heavily influenced by its structural centrality \cite{Nunes_2021}. Crucially, this dynamic hub-like behavior is grounded in recent retrograde tracing data, which revealed that the border region representing the vertical meridian (the TZ) acts as an epicenter of convergence for both dense long-range ipsilateral projections and massive contralateral inputs \cite{Martin_2024}. Being structurally wired as a dense convergence zone, the TZ is anatomically predetermined to act as a highly labile dynamic modulator within the A17 $\rightarrow$ TZ $\rightarrow$ A18 causal network, gating information flow based on the immediate demands of the moving stimulus.


\section{Overcoming the Computational Scale and Open Science}
Translating the theoretical concepts of dynamic connectomics into practical analytical outputs requires overcoming severe computational barriers. Generating stochastic multivariate null models (e.g., surrogate data phase-randomization and configuration models) across continuous sliding windows triggers a combinatorial explosion that typically crashes single-threaded analytical scripts. 
By orchestrating these routines through a Dask-distributed High-Performance Computing (HPC) backend and utilizing `xarray` and `Zarr` for lazy-loading operations, NECTA effectively eliminates the "Big Data" electrophysiology bottleneck. 

Equally vital to this computational leap is the aspiration toward \textit{Open Science} and reproducibility. While systemic barriers remain in widespread adoption, the rigorous parameterization, GitHub/CodeMeta logging, and containerized deployment of NECTA’s codebase represent a significant effort to ensure that our analytical workflows are transparent and globally verifiable. By doing so, it provides the ideal computational environment to validate large-scale biological theories, such as integrating subthreshold oscillatory dynamics into next-generation Spiking Neural Networks (SNNs) for more energy-efficient temporal processing \cite{Lassers2026}. Equally vital to this computational leap is the adherence to \textit{Open Science} and reproducibility. The rigorous parameterization, GitHub/CodeMeta logging, and containerized deployment of NECTA’s codebase ensure that our analytical workflows are transparent and globally verifiable. This pipeline aims to lower the barriers to complex graph-theoretical analysis, striving to allow the scientific community to interactively slice and visualize multidimensional connectomes.

