\chapter{WORKING HYPOTHESIS AND OBJECTIVES}

\section{Working Hypothesis}
Traditional analyses of multichannel electrophysiological data often collapse the temporal dimension, yielding static graphs that obscure the transient coordination of neural ensembles. Furthermore, attempts to extract time-resolved networks are frequently confounded by volume conduction and crippled by the computational intractability of massive stochastic testing.

We hypothesize that integrating phase-based, zero-lag immune connectivity estimators (e.g., Imaginary Coherence and Weighted Phase Lag Index) with dynamically optimized causal models (Partial Directed Coherence) will overcome the artifactual bottlenecks of volume conduction. Critically, to address the profound combinatorial explosion inherent in time-varying analyses, the deployment of a highly scalable, parallel computing architecture—specifically the Dask/Xarray ecosystem—is posited not merely as a software enhancement, but as an indispensable mathematical requirement for the system's survival. The "Big Data" processing scale is a forced consequence of the successive combinatorial routines required for Monte-Carlo simulations, phase-randomized surrogate permutations, and structural null models applied to adaptive topological thresholding over continuous, non-linear dynamic windows.

By deploying this integrated HPC framework, it will be possible to decode the dynamic network evolution of the primary visual cortex. We expect to reveal that the anatomical structural backbone (Areas 17, 18, and the Transition Zone) supports highly flexible, time-varying topological reorganizations—such as transient peaks in Small-Worldness—that align exactly with the active phases of sensory feature binding.

\section{General Objective}
The primary objective of this thesis is to develop, validate, and empirically apply \textbf{NECTA (Neural Electrophysiological Connectivity Time-resolved Analysis)}. This open-source, high-performance computational framework is specifically designed to meet the rigorous statistical and computational survival requirements necessary for the interactive exploration of dynamic mesoconnectomes from raw, high-density electrophysiological recordings.

\section{Specific Objectives}
To achieve the general goal, the project is subdivided into the following specific aims:

\begin{enumerate}
\item \textbf{Data Ingestion and Standardization:} To develop a robust pipeline (the \textit{Isaac} ecosystem) capable of converting dispersed legacy electrophysiological binaries (e.g., SPASS format) into standardized, cloud-native, n-dimensional tensors (Zarr/Xarray) using lazy-loading mechanisms.

\item \textbf{High-Performance Computation Engine:} To architect a distributed pre-computation backend using Dask, capable of orchestrating parallel pipelines for spectral connectivity, phase-based metrics, and graph topology across multiple frequency bands and sliding time windows.

\item \textbf{Advanced Causal Inference:} To implement a self-optimizing Partial Directed Coherence (PDC) module that automatically selects the optimal Multivariate Autoregressive (MVAR) model order via the Bayesian Information Criterion (BIC), thereby preventing overfitting and ensuring rigorous information flow directionality.

\item \textbf{Statistical Rigor and Null Modeling:} To embed strict statistical validation protocols, including False Discovery Rate (FDR) corrections applied to phase-randomized surrogate data, and topological null models (Erdös-Rényi and Configuration Models) to validate graph communities.

\item \textbf{Interactive Visual Analytics:} To construct a reactive, component-based frontend (using Dash and Cytoscape) that translates multidimensional connectivity arrays into an interactive chronologic connectome, allowing frame-by-frame topological inspection.

\item \textbf{Empirical Validation \textit{In Vivo}:} To apply the NECTA framework to an empirical dataset (`c1608a01`) obtained from the feline visual cortex (A17/TZ/A18) to characterize the hierarchical directionality of gamma-band synchronization and the temporal evolution of network efficiency during visual stimulation.
\end{enumerate}