\chapter{INTRODUCTION}

\section{The Primary Visual Cortex and Mesoscopic Networks}
The primary visual cortex has long served as the canonical model for understanding cortical microcircuitry, sensory processing, and network topology. Historically, the architectural and functional parcellation of visual areas—such as Area 17 (A17), Area 18 (A18), and their Transition Zone (TZ) in the feline model—has been extensively mapped using single-unit recordings and static receptive field analyses (Wunderle et al., 2013; Schmidt et al., 2023). These classical approaches have provided the fundamental "structural backbone" of the visual system, detailing how intra- and inter-hemispheric connections form clustered functional communities. However, while static topologies describe the anatomical constraints of neural communication, they obscure the rapid, transient fluctuations in functional integration that are the hallmark of active sensory processing.

\section{From Static Topologies to the ``Dynome''}
Cognitive processes and the integration of complex visual stimuli do not rely on fixed pathways, but rather on highly dynamic, time-varying networks. Recognizing this limitation, the neuroscientific community has begun shifting its focus from static connectomics to the exploration of the "Dynome"—the millisecond-scale evolution of functional networks and their temporal dynamics \cite{kopell2014}. Capturing the Dynome requires shifting the analytical framework from time-averaged correlations to time-resolved state transitions, often achieved through sliding-window approaches \cite{allen2014}. In this context, observing how network metrics, such as \textit{Small-Worldness} and modularity, evolve over brief temporal windows provides a much more accurate representation of the brain's real-time functional state.

\section{Rhythmic Oscillations and Gamma Synchronization}
At the mesoscopic scale, rhythmic neural oscillations—specifically in the Gamma frequency band (30–90 Hz)—act as a fundamental mechanism for cortical communication and information routing. Synchronization in the low-gamma band has been established as a core driver for visual feature binding and functional cohesion across the retinogeniculate and early cortical pathways of the cat \cite{neuenschwander2023}. Furthermore, the directionality of these high-frequency rhythms dictates hierarchical processing. Theoretical and empirical models propose that distinct frequency bands carry specific feedforward or feedback information, establishing temporary master-slave relationships, or "Driver-Receiver" dynamics, between distinct cortical areas \cite{vezoli2021}. Thus, mapping the directionality of gamma synchronization is crucial for deciphering the causal architecture of the visual cortex.

\section{The Methodological Bottleneck in Electrophysiology}
Despite the theoretical necessity of time-resolved, directed connectivity analysis, extracting these networks from dense multi-electrode arrays poses severe methodological and computational challenges. The primary analytical bottleneck in Local Field Potential (LFP) analysis is \textit{volume conduction}—the instantaneous spread of electrical fields through cortical tissue. This phenomenon causes multiple electrodes to record a single biological source simultaneously, generating spurious zero-phase lag correlations that heavily contaminate graph topologies with false-positive edges \cite{bastos2015}.

Furthermore, the transition to high-density electrophysiology (e.g., 32-64 channel matrices) creates a "Big Data" computational crisis. Calculating dynamic, non-linear connectivity metrics (such as Partial Directed Coherence or phase-based indices), generating hundreds of surrogate networks for statistical validation, and tracking graph properties over moving time windows inevitably lead to combinatorial explosions. Traditional, single-threaded analytical scripts are inherently incapable of managing such stochastic, multidimensional loads, often resulting in Out-of-Memory (OOM) failures or prohibitive processing times. 

Consequently, there is a critical need for a modern, high-performance computing (HPC) framework capable of mitigating volume conduction artifacts through advanced phase-based estimators while leveraging distributed parallel processing to interactively render the mesoscopic Dynome.